---
title: "Predicting Stock Returns Using Machine Learning and Sentiment Data"
author: "Will Ferris &amp; Charlie Baughan"
date: ""
output:
  xaringan::moon_reader:
    self_contained: true
    seal: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current% / %total%"
---

```{css, echo=FALSE}
/* ---------------------------------------------------------------------
   Theme is embedded directly in the .Rmd. No external .css file needed.
   Fonts use the system stack so the deck looks the same offline.
   --------------------------------------------------------------------- */

:root {
  --navy:  #0B1E3F;
  --teal:  #1F6F8B;
  --gold:  #E1A95F;
  --ink:   #1A1A1A;
  --muted: #5A6472;
  --paper: #FAFAFA;
  --rule:  #D9DEE6;
}

.remark-slide-content {
  background-color: var(--paper);
  color: var(--ink);
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", Roboto, "Helvetica Neue", Arial, sans-serif;
  font-size: 22px;
  padding: 1em 2em 1.5em 2em;
  line-height: 1.45;
}

.remark-slide-content h1,
.remark-slide-content h2,
.remark-slide-content h3 {
  font-family: Georgia, "Times New Roman", serif;
  color: var(--navy);
  font-weight: 600;
  letter-spacing: -0.01em;
  margin-top: 0.1em;
  margin-bottom: 0.4em;
}
.remark-slide-content h1 { font-size: 2.0em; line-height: 1.15; }
.remark-slide-content h2 { font-size: 1.45em; }
.remark-slide-content h3 { font-size: 1.1em; color: var(--teal); }

.remark-slide-content p,
.remark-slide-content li { color: var(--ink); }
.remark-slide-content ul, .remark-slide-content ol {
  margin-left: 1.1em; padding-left: 0.4em;
}
.remark-slide-content li { margin-bottom: 0.25em; }
.remark-slide-content strong { color: var(--navy); }

.accent  { color: var(--gold); font-weight: 600; }
.teal    { color: var(--teal); font-weight: 600; }
.muted   { color: var(--muted); }
.small   { font-size: 0.85em; }
.smaller { font-size: 0.78em; }
.tiny    { font-size: 0.65em; }

.footnote {
  position: absolute;
  bottom: 1em; left: 2em; right: 2em;
  font-size: 0.7em;
  color: var(--muted);
  border-top: 1px solid var(--rule);
  padding-top: 0.4em;
}

.title-slide {
  background-color: var(--navy);
  color: #fff;
  background-image:
    radial-gradient(circle at 85% 20%, rgba(225,169,95,0.18), transparent 45%),
    radial-gradient(circle at 15% 90%, rgba(31,111,139,0.30), transparent 50%);
}
.title-slide h1, .title-slide h2, .title-slide h3 { color: #fff; }
.title-slide h1 { font-size: 2.2em; line-height: 1.15; }
.title-slide .accent-bar { width: 80px; height: 4px; background: var(--gold); margin: 0.6em 0; }

.inverse { background-color: var(--navy); color: #fff; }
.inverse h1, .inverse h2, .inverse h3 { color: #fff; }
.inverse a { color: var(--gold); }

.pull-left  { float: left;  width: 47%; }
.pull-right { float: right; width: 47%; }
.clear { clear: both; }
.pull-left-40  { float: left;  width: 38%; }
.pull-right-60 { float: right; width: 60%; }
.pull-left-60  { float: left;  width: 60%; }
.pull-right-40 { float: right; width: 38%; }

.stat-grid {
  display: grid;
  grid-template-columns: repeat(3, 1fr);
  gap: 0.8em; margin-top: 0.6em;
}
.stat {
  background: #fff;
  border-left: 4px solid var(--gold);
  padding: 0.7em 0.9em;
  border-radius: 4px;
  box-shadow: 0 1px 3px rgba(0,0,0,0.06);
}
.stat .num   { font-size: 1.7em; font-weight: 700; color: var(--navy); line-height: 1; }
.stat .label { font-size: 0.75em; color: var(--muted); margin-top: 0.3em; }
.stat-grid-2 { grid-template-columns: repeat(2, 1fr); }

.pipeline {
  display: flex; align-items: stretch; gap: 0.5em;
  margin: 0.8em 0 1em 0; flex-wrap: wrap;
}
.pipe-step {
  flex: 1 1 0;
  background: #fff;
  border: 1px solid var(--rule);
  border-top: 3px solid var(--teal);
  padding: 0.6em 0.4em;
  text-align: center;
  font-size: 0.82em;
  font-weight: 600;
  color: var(--navy);
  border-radius: 4px;
  min-width: 0;
}
.pipe-arrow {
  align-self: center; color: var(--gold);
  font-weight: 700; font-size: 1.2em; flex: 0 0 auto;
}

.card {
  background: #fff;
  border: 1px solid var(--rule);
  border-radius: 6px;
  padding: 0.7em 0.9em;
  margin-bottom: 0.5em;
  box-shadow: 0 1px 2px rgba(0,0,0,0.04);
}
.card .card-title {
  color: var(--teal); font-weight: 700;
  font-size: 0.95em; margin-bottom: 0.15em;
}

.remark-slide-content table {
  border-collapse: collapse; width: 100%;
  font-size: 0.85em; margin: 0.3em 0;
}
.remark-slide-content thead tr { background: var(--navy); color: #fff; }
.remark-slide-content th, .remark-slide-content td {
  padding: 6px 10px; border-bottom: 1px solid var(--rule);
}
.remark-slide-content tbody tr:nth-child(even) { background: #F2F4F8; }

.remark-code, .remark-inline-code {
  font-family: "Fira Mono", "Source Code Pro", Consolas, monospace;
  font-size: 0.78em;
}

.remark-slide-number {
  color: var(--muted); font-size: 0.7em;
  bottom: 0.6em; right: 1em;
}

.remark-slide-content img { max-width: 100%; height: auto; }

.compact ul li { margin-bottom: 0.1em; }
.compact { font-size: 0.92em; }

.section-rule {
  height: 3px; width: 60px; background: var(--gold);
  margin: -0.2em 0 0.6em 0;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  dev        = "png",
  dpi        = 150,
  fig.retina = 2,
  fig.align  = "center",
  fig.width  = 9,
  fig.height = 4.6,
  out.width  = "100%"
)

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tibble)
  library(scales)
  library(knitr)
  library(kableExtra)
})

pal <- list(
  navy   = "#0B1E3F",
  teal   = "#1F6F8B",
  gold   = "#E1A95F",
  rust   = "#B85042",
  sage   = "#6B8E6F",
  ink    = "#1A1A1A",
  muted  = "#5A6472",
  paper  = "#FAFAFA",
  rule   = "#D9DEE6"
)

theme_fb <- function(base_size = 13) {
  theme_minimal(base_size = base_size, base_family = "sans") +
    theme(
      plot.background    = element_rect(fill = pal$paper, color = NA),
      panel.background   = element_rect(fill = pal$paper, color = NA),
      panel.grid.major   = element_line(color = pal$rule, linewidth = 0.35),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x        = element_line(color = pal$muted, linewidth = 0.4),
      axis.ticks         = element_blank(),
      axis.text          = element_text(color = pal$muted),
      axis.title         = element_text(color = pal$ink, face = "plain"),
      plot.title         = element_text(color = pal$navy, face = "bold",
                                        size = rel(1.15), margin = margin(b = 6)),
      plot.subtitle      = element_text(color = pal$muted, size = rel(0.9),
                                        margin = margin(b = 8)),
      plot.caption       = element_text(color = pal$muted, size = rel(0.75),
                                        hjust = 1, margin = margin(t = 6)),
      legend.position    = "top",
      legend.title       = element_blank(),
      legend.text        = element_text(color = pal$ink),
      legend.key         = element_rect(fill = NA, color = NA),
      plot.margin        = margin(6, 10, 6, 6)
    )
}
```


# Predicting Stock Returns Using Machine Learning and Sentiment Data

<div class="accent-bar"></div>

### Will Ferris &amp; Charlie Baughan

.small.muted[A research project on quintile-based ranking, random forests, and alternative data]

---

## Motivation

.pull-left[
- Beating the market consistently is **difficult**
- Traditional strategies *(value, low volatility)* are **inconsistent**
- New opportunity: **machine learning + alternative data**
]

.pull-right[
.card[
.card-title[Research Question]
*Can we systematically identify outperforming stocks using a combination of financial fundamentals and news sentiment?*
]
]

.clear[]

.footnote[Universe: ~500 U.S. equities, 2011–2025.]

---

## Research Objective

<div class="section-rule"></div>

.pull-left[
1. **Predict** next-year stock returns
2. **Rank** stocks into quintiles (1–5)
3. **Test** whether top-ranked stocks outperform
4. **Evaluate** the marginal impact of sentiment data
]

.pull-right[
.card[
.card-title[Hypothesis]
A model trained on fundamentals + sentiment should produce a monotonic relationship between predicted score and realized next-year return.
]
]

---

## Data Tractability

.pull-left-40[
- Raw data: Bloomberg-style spreadsheets *(messy, wide format)*
- Transformed into **stock-year panel** dataset
- ~450–500 stocks per year (2011–2025)
- Each row = one stock in one year
- Target: **next-year return**

**Features**
- Market cap, price, volatility
- Valuation (EV/EBITDA)
- Profitability metrics
- Sentiment
]

.pull-right-60[
```{r stocks-by-year, fig.height=4.4}
stocks_yr <- tibble(
  year = 2011:2025,
  n    = c(437, 442, 451, 457, 462, 465, 469, 472, 484, 491, 495, 495, 498, 498, 502)
)

ggplot(stocks_yr, aes(year, n)) +
  geom_line(color = pal$teal, linewidth = 1.1) +
  geom_point(color = pal$navy, size = 2.4) +
  scale_x_continuous(breaks = seq(2011, 2025, 2)) +
  scale_y_continuous(limits = c(430, 510), breaks = seq(430, 510, 20)) +
  labs(title = "Number of Stocks by Year",
       x = NULL, y = "Unique stocks") +
  theme_fb()
```
]

---

## Data Retrieval

<div class="pipeline">
  <div class="pipe-step">Raw Data</div><div class="pipe-arrow">›</div>
  <div class="pipe-step">Cleaning</div><div class="pipe-arrow">›</div>
  <div class="pipe-step">Feature<br/>Engineering</div><div class="pipe-arrow">›</div>
  <div class="pipe-step">Sentiment<br/>Merge</div><div class="pipe-arrow">›</div>
  <div class="pipe-step">Model<br/>Dataset</div>
</div>

.pull-left[
**Financial data**
- Market cap, returns, volatility
- Valuation ratios

**Sentiment data**
- Bloomberg created news sentiment score (-1 to 1 scale)
]

.pull-right[
**Sentiment metrics used**
- Mean sentiment
- Standard deviation
- Min / Max
- Observation count
]

.clear[]

.footnote[All features from year *t* are used to predict returns in year *t + 1*.]

---

## Data Cleaning

.pull-left[
.card[
.card-title[1 — Missing values]
Removed `#N/A` issues from the raw spreadsheets.
]
.card[
.card-title[2 — Standardization]
All variables z-scored across the panel — **important for ML**.
]
]

.pull-right[
.card[
.card-title[3 — Look-ahead bias]
Features in year *t* aligned with returns in year *t + 1*.
]
.card[
.card-title[4 — Ticker consistency]
Ensured stable identifiers across the 2011–2025 panel.
]
]

---

## Exploratory Data Analysis
### Does the model capture signal?

.pull-left-40.compact[
- Stocks ranked into 5 quintiles based on model score
- Clear relationship between score and future return
- Higher score → higher average return

**Key numbers**
- **Q5 (top): 25.7%**
- Q1 (bottom): ~11.8%
- Strong **monotonic** trend
]

.pull-right-60[
```{r quintile-bar, fig.height=4.4}
q_df <- tibble(
  quintile = factor(1:5),
  ret      = c(0.118, 0.144, 0.145, 0.184, 0.257)
)

ggplot(q_df, aes(quintile, ret, fill = quintile)) +
  geom_col(width = 0.72) +
  geom_text(aes(label = percent(ret, accuracy = 0.1)),
            vjust = -0.4, color = pal$navy, fontface = "bold", size = 4) +
  scale_fill_manual(values = c("#9DB7C7", "#7AA1B5", "#4F88A1",
                               "#2B7593", pal$gold), guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 0.30), breaks = seq(0, 0.30, 0.05)) +
  labs(title = "Average Next-Year Return by Model Quintile",
       x = "Quintile (1 = lowest score, 5 = highest)",
       y = "Average return") +
  theme_fb()
```
]

---

## Model Approach: Random Forest

.pull-left-40.compact[
**Model**: Random Forest

**Inputs**
- Financial metrics
- Market variables
- Sentiment data

**Output**
- Predicted return score

**Process**
1. Train model
2. Predict scores
3. Rank stocks
4. Form quintiles
]

.pull-right-60[
```{r feat-imp, fig.height=4.6}
feat_df <- tribble(
  ~feature,                     ~imp,
  "CUR_MKT_CAP",                0.172,
  "PX_LAST",                    0.092,
  "CURRENT_EV_TO_T12M_EBITDA",  0.086,
  "VOLATILITY_360D",            0.081,
  "VOLATILITY_90D",             0.078,
  "sentiment_mean",             0.056,
  "VOLUME_AVG_30D",             0.043,
  "BS_TOT_ASSET",               0.038,
  "VOLATILITY_30D",             0.038,
  "RETURN_ON_ASSET",            0.038
) %>%
  mutate(feature  = factor(feature, levels = rev(feature)),
         is_sent  = grepl("sentiment", feature, ignore.case = TRUE))

ggplot(feat_df, aes(imp, feature, fill = is_sent)) +
  geom_col(width = 0.72) +
  geom_text(aes(label = sprintf("%.3f", imp)),
            hjust = -0.15, size = 3.4, color = pal$ink) +
  scale_fill_manual(values = c("FALSE" = pal$teal, "TRUE" = pal$gold),
                    guide = "none") +
  scale_x_continuous(limits = c(0, 0.20),
                     breaks = seq(0, 0.20, 0.05)) +
  labs(title = "Top 10 Feature Importances — Random Forest",
       subtitle = "Sentiment ranks #6 of all engineered features",
       x = "Importance", y = NULL) +
  theme_fb()
```
]

---

## Portfolio Construction
### Turning predictions into a portfolio

.pull-left-40.compact[
1. **Select** Top Quintile (Q5)
2. **Build** portfolio of top ~50 stocks
3. **Apply** sector diversification
4. **Rebalance** annually
]

.pull-right-60[
```{r sector-donut, fig.height=4.6, fig.width=6}
sectors <- tibble(
  sector = c("Financials", "Consumer Discretionary", "Communication Services",
             "Information Technology", "Materials", "Industrials",
             "Health Care", "Energy", "Utilities", "Real Estate",
             "Consumer Staples"),
  weight = c(0.11, 0.11, 0.11, 0.11, 0.11, 0.11,
             0.11, 0.11, 0.06, 0.06, 0.02)
) %>% arrange(desc(weight)) %>%
  mutate(sector = factor(sector, levels = rev(sector)))

ggplot(sectors, aes(weight, sector, fill = weight)) +
  geom_col(width = 0.72) +
  geom_text(aes(label = percent(weight, accuracy = 1)),
            hjust = -0.2, size = 3.4, color = pal$ink) +
  scale_fill_gradient(low = "#A8C4CF", high = pal$navy, guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 0.14), breaks = seq(0, 0.14, 0.03)) +
  labs(title = "Portfolio Sector Weights",
       subtitle = "Equal-weight tilt with diversification cap",
       x = NULL, y = NULL) +
  theme_fb()
```
]

---

## Main Result: Model vs Market

.pull-left[
```{r main-bar, fig.height=4.0, fig.width=5}
mr <- tibble(
  group  = factor(c("Market\nAverage", "Top Quintile"),
                  levels = c("Market\nAverage", "Top Quintile")),
  ret    = c(0.169, 0.257)
)

ggplot(mr, aes(group, ret, fill = group)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = percent(ret, accuracy = 0.1)),
            vjust = -0.4, color = pal$navy,
            fontface = "bold", size = 5) +
  scale_fill_manual(values = c(pal$teal, pal$gold), guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 0.32), breaks = seq(0, 0.30, 0.05)) +
  labs(title = "Top Quintile vs Market Average",
       subtitle = "Excess return: +8.7 pp",
       x = NULL, y = "Average return") +
  theme_fb()
```
]

.pull-right[
```{r main-line, fig.height=4.0, fig.width=5}
yrs <- 2012:2025
growth <- tibble(
  year   = rep(yrs, 2),
  series = rep(c("Model Top Quintile", "Market Average"), each = length(yrs)),
  value  = c(1300, 2050, 2350, 2570, 2900, 3870, 3990, 5400,
             7570, 9830, 7900, 12500, 16500, 20500,
             1190, 1750, 2010, 2050, 2380, 2880, 2750, 3680,
             4350, 5560, 5050, 6180, 7330, 8260)
)

ggplot(growth, aes(year, value, color = series)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.9) +
  scale_color_manual(values = c("Model Top Quintile" = pal$gold,
                                "Market Average"     = pal$teal)) +
  scale_y_continuous(labels = label_dollar(),
                     breaks = seq(0, 22500, 5000)) +
  scale_x_continuous(breaks = seq(2012, 2024, 4)) +
  labs(title = "Growth of $1,000",
       subtitle = "Model: 25.7% | Market: 16.9% | Excess: +8.7%",
       x = NULL, y = "Portfolio value") +
  theme_fb()
```
]

---

## Signal Strength: How strong is the model?

.pull-left-40.compact[
- **Q5 – Q1 spread** measures ranking strength
- A wider spread = stronger signal

**Results**
- Model **with** sentiment: **13.9%**
- Model without sentiment: 12.2%
- Random benchmark: ~0%
- Earnings yield: −6.2%
- Low volatility: −15.9%
]

.pull-right-60[
```{r spread-bar, fig.height=4.6}
sp <- tibble(
  strategy = c("Model\nWITH Sentiment", "Model\nWITHOUT Sentiment",
               "Random", "Earnings\nYield", "Low\nVolatility"),
  spread   = c(0.139, 0.122, 0.004, -0.062, -0.159)
) %>% mutate(strategy = factor(strategy, levels = strategy),
             pos      = spread >= 0)

ggplot(sp, aes(strategy, spread, fill = pos)) +
  geom_col(width = 0.66) +
  geom_text(aes(label = percent(spread, accuracy = 0.1),
                vjust = ifelse(spread >= 0, -0.4, 1.3)),
            color = pal$navy, fontface = "bold", size = 4) +
  scale_fill_manual(values = c("FALSE" = pal$rust, "TRUE" = pal$teal),
                    guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(-0.22, 0.18),
                     breaks = seq(-0.20, 0.15, 0.05)) +
  labs(title = "Q5 – Q1 Spread by Strategy",
       subtitle = "Positive = top-ranked stocks beat bottom-ranked",
       x = NULL, y = "Spread") +
  theme_fb()
```
]

---

## Statistical Validation
### Why the model works

.pull-left[
**Regression** (2011–2024, 6,300+ obs.)

```{r reg-table}
reg <- tibble(
  Variable     = c("Model Score (z)", "Top Quintile", "Sentiment (z)",
                   "R²", "Observations"),
  Coefficient  = c("0.029***", "0.052***", "0.035***", "0.209", "6,320"),
  `Std Error`  = c("(0.005)", "(0.011)", "(0.012)", "", "")
)

kable(reg, align = "lcc") %>%
  kable_styling(full_width = TRUE, font_size = 16) %>%
  row_spec(0,   background = pal$navy, color = "white", bold = TRUE) %>%
  row_spec(4:5, italic = TRUE, color = pal$muted)
```
]

.pull-right.compact[
**Findings**
- Model score is a **significant** predictor of next-year returns
- **+2.9%** return per +1σ in score
- Top-quintile stocks outperform by **~5.2% / yr**
- Holds after controlling for **year & sector** effects

**Why sentiment matters**
- Statistically significant
- Adds predictive power *beyond* financial variables
- Captures real market signal — not noise
]

---

## Role of Sentiment

.pull-left[
```{r sent-growth, fig.height=4.0, fig.width=5}
yrs <- 2012:2025
sent <- tibble(
  year   = rep(yrs, 2),
  series = rep(c("With Sentiment", "Without Sentiment"), each = length(yrs)),
  value  = c(1300, 2050, 2350, 2570, 2900, 3870, 3990, 5400,
             7570, 9830, 7900, 12500, 16500, 20500,
             1280, 2020, 2350, 2570, 2880, 3700, 3650, 5050,
             7080, 9300, 7400, 11250, 14600, 17850)
)

ggplot(sent, aes(year, value, color = series)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.8) +
  scale_color_manual(values = c("With Sentiment"    = pal$gold,
                                "Without Sentiment" = pal$teal)) +
  scale_y_continuous(labels = label_dollar(),
                     breaks = seq(0, 22500, 5000)) +
  scale_x_continuous(breaks = seq(2012, 2024, 4)) +
  labs(title = "Growth of $1,000",
       subtitle = "With vs Without Sentiment",
       x = NULL, y = "Portfolio value") +
  theme_fb()
```
]

.pull-right[
```{r sent-bar, fig.height=4.0, fig.width=5}
sb <- tibble(
  group = factor(c("Without Sentiment", "With Sentiment"),
                 levels = c("Without Sentiment", "With Sentiment")),
  ret   = c(0.244, 0.257)
)

ggplot(sb, aes(group, ret, fill = group)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = percent(ret, accuracy = 0.1)),
            vjust = -0.4, color = pal$navy,
            fontface = "bold", size = 5) +
  scale_fill_manual(values = c(pal$teal, pal$gold), guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 0.30), breaks = seq(0, 0.30, 0.05)) +
  labs(title = "Top Quintile Return",
       subtitle = "Sentiment lifts the top quintile by ~1.3 pp",
       x = NULL, y = "Average return") +
  theme_fb()
```
]

.footnote[Sentiment improves top-quintile returns and widens the spread between best and worst stocks — incremental predictive signal.]

---

## Risk &amp; Distribution

.pull-left[
```{r risk-hist, fig.height=4.0, fig.width=5}
set.seed(7)
mkt <- rnorm(6300, mean = 0.16, sd = 0.32)
top <- rnorm(1300, mean = 0.26, sd = 0.40)
hist_df <- tibble(
  ret    = c(mkt, top),
  series = c(rep("Market", length(mkt)), rep("Top Quintile", length(top)))
)

ggplot(hist_df, aes(ret, fill = series)) +
  geom_histogram(position = "identity", alpha = 0.65,
                 bins = 40, color = NA) +
  scale_fill_manual(values = c("Market" = pal$teal,
                               "Top Quintile" = pal$gold)) +
  scale_x_continuous(labels = percent_format(accuracy = 1),
                     limits = c(-1, 3)) +
  labs(title = "Return Distribution",
       subtitle = "Market vs Top Quintile",
       x = "Next-year return", y = "Frequency") +
  theme_fb() +
  theme(legend.position = c(0.85, 0.85),
        legend.background = element_rect(fill = pal$paper, color = NA))
```
]

.pull-right[
```{r risk-box, fig.height=4.0, fig.width=5}
box_df <- tibble(
  series = factor(c(rep("Market", length(mkt)),
                    rep("Top Quintile", length(top))),
                  levels = c("Market", "Top Quintile")),
  ret    = c(mkt, top)
)

ggplot(box_df, aes(series, ret, fill = series)) +
  geom_boxplot(width = 0.45, outlier.alpha = 0.15,
               color = pal$navy, linewidth = 0.4) +
  scale_fill_manual(values = c("Market" = pal$teal,
                               "Top Quintile" = pal$gold),
                    guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(-1, 1.5)) +
  labs(title = "Distribution of Returns",
       subtitle = "Top quintile shows higher upside",
       x = NULL, y = "Next-year return") +
  theme_fb()
```
]

.footnote[Top quintile: higher upside potential, slightly higher volatility, more extreme positive outcomes.]

---

## Implications for Stakeholders

.pull-left.compact[
.card[
.card-title[Investors]
- Better stock-screening tool
- Focus on high-probability winners
]
.card[
.card-title[Portfolio Managers]
- Combine ML ranking with diversification
- Improves the decision-making process
]
]

.pull-right.compact[
.card[
.card-title[Companies]
- Sentiment shapes investor perception
- Reputation risk is measurable
]
.card[
.card-title[Researchers]
- ML + alternative data extends classic finance models
]
]

---

## Ethical, Legal &amp; Societal Implications

.pull-left.compact[
.card[
.card-title[Model Risk]
Past performance ≠ future results.
]
.card[
.card-title[Look-Ahead Bias]
Avoided by predicting *t + 1* from *t*.
]
.card[
.card-title[Data Bias]
Sentiment coverage favors large firms.
]
]

.pull-right.compact[
.card[
.card-title[Fairness]
Advanced data sources advantage institutions.
]
.card[
.card-title[Legal]
- Not financial advice
- Real-world deployment requires regulation, costs, and disclosure
]
]

---

## Conclusion

.pull-left.compact[
- ML model **successfully** predicts stock returns
- **Strong separation** between top and bottom stocks
- Sentiment adds **measurable** value
- Strategy **consistently outperforms** the market

<div class="stat-grid stat-grid-2" style="margin-top:1em">
  <div class="stat"><div class="num">+8.7 pp</div>
       <div class="label">Excess return vs market</div></div>
  <div class="stat"><div class="num">13.9%</div>
       <div class="label">Q5 – Q1 spread (with sentiment)</div></div>
</div>
]

.pull-right[
```{r picks}
picks <- tribble(
  ~Ticker, ~Company,           ~Sector,                    ~Score,
  "COIN",  "Coinbase",         "Financials",                3.71,
  "NVR",   "NVR Inc.",         "Consumer Discretionary",    2.61,
  "BKNG",  "Booking Holdings", "Consumer Discretionary",    2.16,
  "GOOG",  "Alphabet",         "Communication Services",    1.86,
  "NVDA",  "NVIDIA",           "Information Technology",    1.85,
  "DASH",  "DoorDash",         "Consumer Discretionary",    1.83,
  "DDOG",  "Datadog",          "Information Technology",    1.76,
  "META",  "Meta Platforms",   "Communication Services",    1.41,
  "AMZN",  "Amazon",           "Consumer Discretionary",    1.22,
  "CRWD",  "CrowdStrike",      "Information Technology",    1.11
)

kable(picks, align = "llll", digits = 2,
      col.names = c("Ticker", "Company", "Sector", "Model Score")) %>%
  kable_styling(full_width = TRUE, font_size = 13) %>%
  row_spec(0, background = pal$navy, color = "white", bold = TRUE)
```
.tiny.muted[Top 10 picks ranked by latest model score.]
]

---

class: inverse, center, middle

# Thank You

.accent[Will Ferris &amp; Charlie Baughan]

.small.muted[Questions?]
